---
title: TAPS Economic Questions
author: JBrandon Duck-Mayr
output:
    md_document:
        variant: gfm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
quack::attach_common_packages()
options(readr.show_col_types = FALSE)
```

# TAPS Economic Questions

From December 2011 to January 2018, the [Weidenbaum Center on the Economy, Government, and Public Policy](https://wc.wustl.edu/) at [Washington University in St. Louis](https://wustl.edu/) ran [The American Panel Survey](https://wc.wustl.edu/american-panel-survey) (TAPS), a monthly survey revisiting the same nationally representative panel of respondents.
Among the recurring sets of questions asked of respondents were the following questions about the economy (with observed answer choices underneath each question):

- ECON1A: [in your household] Are the economic conditions in your household and in the countryare getting better, worse, or not changing much?
  - Refused (coded in the datasets here as `NA`)
  - Getting much better (coded in the datasets here as 5)
  - Getting somewhat better (coded in the datasets here as 4)
  - Not changing much (coded in the datasets here as 3)
  - Getting somewhat worse (coded in the datasets here as 2)
  - Getting much worse (coded in the datasets here as 1)
  - Not sure (coded in the datasets here as `NA`)
- ECON1B: [In the country as a whole] Are the economic conditions in your household and in the countryare getting better, worse, or not changing much?
  - Refused (coded in the datasets here as `NA`)
  - Getting much better (coded in the datasets here as 5)
  - Getting somewhat better (coded in the datasets here as 4)
  - Not changing much (coded in the datasets here as 3)
  - Getting somewhat worse (coded in the datasets here as 2)
  - Getting much worse (coded in the datasets here as 1)
  - Not sure (coded in the datasets here as `NA`)
- ECON2A: [In your household] Are the current economic conditions in your household and in the country excellent, good, only fair, or poor?
  - Refused (coded in the datasets here as `NA`)
  - Excellent (coded in the datasets here as 4)
  - Good (coded in the datasets here as 3)
  - Only fair (coded in the datasets here as 2)
  - Poor (coded in the datasets here as 1)
  - Not sure (coded in the datasets here as `NA`)
- ECON2B: [In the country as a whole] Are the current economic conditions in your household and in the country excellent, good, only fair, or poor?
  - Refused (coded in the datasets here as `NA`)
  - Excellent (coded in the datasets here as 4)
  - Good (coded in the datasets here as 3)
  - Only fair (coded in the datasets here as 2)
  - Poor (coded in the datasets here as 1)
  - Not sure (coded in the datasets here as `NA`)
- ECON4A: [I expect to spend...] Compared to what you have done in the past few months, do you expect to spend or save
  - Refused (coded in the datasets here as `NA`)
  - Much more (coded in the datasets here as 5)
  - More (coded in the datasets here as 4)
  - About the same (coded in the datasets here as 3)
  - Less (coded in the datasets here as 2)
  - Much less (coded in the datasets here as 1)
- ECON4B: [I expect to save...] Compared to what you have done in the past few months, do you expect to spend or save
  - Refused (coded in the datasets here as `NA`)
  - Much more (coded in the datasets here as 5)
  - More (coded in the datasets here as 4)
  - About the same (coded in the datasets here as 3)
  - Less (coded in the datasets here as 2)
  - Much less (coded in the datasets here as 1)

The TAPS data yearly releases are available at <https://wc.wustl.edu/taps-data-archive>.
I downloaded these releases and named the files "TAPSdataYYYY.sav", where YYYY is replaced with the relevant year, then used the code in `pull-econ-questions.R` to combine, recode, and reshape the data in ways that may be useful.
(I did not include those `.sav` files in this repository as collectively they are about 80 Mb).
This repository now includes the responses to those questions in two formats:

```{r read-in-data}
long_data = read_csv("TAPS-econ-data-long.csv")
wide_data = read_csv("TAPS-econ-data-wide.csv")
```

- `TAPS-econ-data-long.csv`
  + `r ncol(long_data)` columns
    - `WUSTLID` gives the unique respondent ID
    - `wave` indicates what wave the responses in each row respond to
    - One column for each of the 6 questions described above
  + `r nrow(long_data)` rows
    - Each row corresponds to all answers to the economic questions given by a respondent in a particular wave of TAPS
- `TAPS-econ-data-wide.csv`
  + `r ncol(wide_data)` columns
    - `wave` indicates what wave the responses in each row respond to
    - `question` indicates which of the 6 questions described above the responses in each row respond to
    - One column for each of the `r length(unique(long_data$WUSTLID))` TAPS respondents (the column name is the respondent's `WUSTLID`)
  + `r nrow(wide_data)` rows
    - Each row corresponds to all answers given to a particular economic question in a particular wave of TAPS
